### 2020-02-01 background.tex
more small changes, still just starting those two sections

\paragraph{Statistical mechanical theory of NNs.}

Statistical mechanical ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
our best-performing metrics (i.e., those using fitted PL exponents and HT-SR Theory) are based on statistical mechanics ideas~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}.
However, that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
We will \emph{use} statistical mechanics theory more as it is used in quantitative finance.
Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.

>>>>

\paragraph{Statistical mechanics theory for NNs.}

Statistical mechanics ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.  
However, we emphasize that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
We will \emph{use} statistical mechanics theory more as it is used in quantitative finance.
Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.

### 2020-02-01 cv_models.tex
more small changes, still just starting those two sections

\begin{table}[t]
\small
\begin{center}
%\begin{tabular}{|p{1in}|c|c|c|c|c|}
\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
%\begin{tabular}{|l|c|c|c|c|c|}
\hline
%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 Series & Models & Metric                     & Metric                     & Metric         & Metric                                   \\
\hline
 VGG       &  6 & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
 ResNet    &  5 & 0.9  & 1.4  & \textbf{0.61} & 0.66           \\
 ResNet-1K & 19 & 2.4  & 3.6  & \textbf{1.8}  & 1.9            \\
 DenseNet  &  4 & 0.3  & 0.26 & \textbf{0.16} & 0.21           \\
\hline
\end{tabular}
\end{center}
\caption{RMSE (smaller is better) for linear fits of quality metrics to Reported Top1 Test Error for pretrained models in each architecture series.  VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K. 
\michael{Are the numbers in this table stale; if they get updated, then we need to modify the test at at least two places, on actual numbers and on what is best.}
}
\label{table:cv-models}
\end{table}

>>>>

\begin{table}[t]
\small
\begin{center}
%\begin{tabular}{|p{1in}|c|c|c|c|c|}
\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
%\begin{tabular}{|l|c|c|c|c|c|}
\hline
%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
% Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
        &        & Log                   & Log                        & Weighted       & Log                                 \\
        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
 Series & Models & Metric                & Metric                     & Metric         & Metric                              \\
\hline
 VGG       &  6 & 0.56 & 0.53 & 0.48          & \textbf{0.42}  \\
 ResNet    &  5 & 0.9  & 1.4  & \textbf{0.61} & 0.66           \\
 ResNet-1K & 19 & 2.4  & 3.6  & \textbf{1.8}  & 1.9            \\
 DenseNet  &  4 & 0.3  & 0.26 & \textbf{0.16} & 0.21           \\
\hline
\end{tabular}
\end{center}
\caption{RMSE (smaller is better) for linear fits of quality metrics to Reported Top1 Test Error for pretrained models in each architecture series.  VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K. 
\michael{Are the numbers in this table stale; if they get updated, then we need to modify the test at at least two places, on actual numbers and on what is best.}
}
\label{table:cv-models}
\end{table}

### 2020-02-01 introduction.tex
more small changes, still just starting those two sections

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo}.
\michael{LINK TO: Google Colab, IS THIS THAT REPO.}

>>>>

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
with fully reproducible Jupyter notebooks.%
\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}}
\michael{LINK TO: Google Colab, IS THIS THAT REPO.}
\michael{FIX THIS PAR.}

### 2020-02-01 methods.tex
more small changes, still just starting those two sections

Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
%
We have examined a large number of possible quality metrics.
The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
Consider the following.
\begin{itemize}
\item 
Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
\item 
Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
\michael{Should this be $\sqrt{\lambda_{max}}$?}
\item 
Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
\item 
$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
\end{itemize}
The first two quantities are well-known in ML.
The last two deserve special mention.
For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
$ %% $$
\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
$ %% $$
and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
\end{equation}
where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.

For norm-based metrics, we consider the average log norm to the appropriate power.
For example, for the \emph{$\alpha$-Norm metric}, 
\begin{equation}
\label{eqn:sum_log_alpha_norm_alpha}
\sum\nolimits_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l}^{\alpha_l} 
=
\sum\nolimits_l \alpha_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l} .
\end{equation}
Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log Complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  ,
\end{equation}
 takes the form of an Average Log Norm.
For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
(When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.)
For the Weighted Alpha Norm, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.

@@

\paragraph{Normalization of Empirical Matrices.}  
Normalization is an important, if underappreciated, practical issue.
Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
\footnote{For formal proofs of Heavy Tailed results, one typically needs a different normalization such as $1/N^{1-\alpha}$.}
Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.

>>>>

Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
%
We have examined a large number of possible quality metrics.
The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
Consider the following.
\begin{itemize}
\item 
Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
\item 
Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
\michael{Should this be $\sqrt{\lambda_{max}}$?}
\item 
Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
\item 
$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
\end{itemize}
The first two quantities are well-known in ML.
The last two deserve special mention.
For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
$ %% $$
\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
$ %% $$
and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
\end{equation}
where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.

For norm-based metrics, we use the average of the log norm to the appropriate power.
For example, consider the \emph{$\alpha$-Norm metric}, 
\begin{equation}
\label{eqn:sum_log_alpha_norm_alpha}
\sum\nolimits_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l}^{\alpha_l} 
=
\sum\nolimits_l \alpha_l \log \Vert\mathbf{W}_l\Vert_{\alpha_l} .
\end{equation}
Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log Complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  ,
\end{equation}
 takes the form of an Average Log Norm.
For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly. 
(When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.)
For the Weighted Alpha Norm, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it cannot be taken out of the sum.

@@

\paragraph{Normalization of Empirical Matrices.}  
Normalization is an important, if underappreciated, practical issue.
Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
\footnote{For formal proofs of Heavy Tailed results, one typically needs a normalization such as $1/N^{1-\alpha}$.}
Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.

### 2020-02-01 nlp_models.tex
more small changes, still just starting those two sections

In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several Natural Language Processing (NLP) model architectures.
NLP models are different in many ways than the CV models we considered in Section~\ref{sxn:cv}, and thus they should provide a very different test of norm based metrics and HT-SR metrics than do CV models.

In particular, within the past two years, nearly 100 open source, pre-trained DNNs for NLP have emerged, based on the revolutionary Transformer architecture.
This includes variants of BERT, Transformer-XML, GPT, etc.
The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
In contrast to the smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger, 
and frequently under-correlated, with generally larger Heavy Tailed Power Law exponents $\alpha$ (when fitting the ESD).

Nevertheless, norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.

Here, PL-based metrics 
(the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $)

we breifly look at a few popular pretrained NLP DNNs to demonstrate that HT-SR Theory 
also
applies to NLP models and not just CV models, and to highlight how to use the theory to identify
poorly trained models where the empirical norm metrics will perform poorly.

\michael{We show that this not only works for CV, so we look at NLP.
We now look in detail at the alpha metric, the spectral norm, and the new norm.
We compare good and bad models, and a series of increasingly big models
trained on big data sets.  We show why alpha works but spectral norm is not enough

}

\michael{Need to say that such large $\alpha$ values don't mean much.}

\michael{Need to highlight difference between $\alpha$ and $\hat{\alpha}$.}

apply more generally, where they differ, e.g., how the layer Spectral Norm behaves unexpectedly, while the $\alpha$-based metrics can give insight into how well trained a model is.

@@

\charles{More details here.}
The GPT models we analyze are deployed with the popular HuggingFace PyTorch library.
GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
The projection layer concatenates the self-attention results into a vector (of dimension $768$).
This gives $50$ large matrices.
\michael{Fix that.}

Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
GPT  has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively. 
Interstingly, they also have very spectral properties, also shown below. XXX.  WHICH.  MOVE.
\michael{Fix that.}

The OpenAI GPT2 (English) models are: \nred{gpt-small, gpt-medium, gpt-large, and gpt-xl}, \
having include $12, 24, 36, \text{and }48$ layers, resp., with increasingly larger weight matrices.
The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, using \emph{weightwatcher (0.2.7)},
and with fully reproducible Jupyter notebooks.\footnote{\url{https://github.com/CalculatedContent/kdd2020}}
\michael{Fix that.}

@@

\begin{table}[t]
\small
\begin{center}
%\begin{tabular}{|p{1in}|c|c|c|c|c|}
\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
\hline
%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
% Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
        & Num.   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
 Series & Layers & Metric                     & Metric                          & Metric         & Metric                                   \\
\hline
GPT & 49 & 1.64  & 1.72 & 7.01 & 7.28 \\
GPT (small) & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
GPT2 large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
GPT2 xl & 194 & 1.86 & 1.92 & 7.17 & 7.51 \\
\hline
\end{tabular}
\end{center}
\caption{Average value for each of the four metrics for pretrainnd OpenAI GPT and GPT2 models.}
\label{table:nlp}
\end{table}

>>>>

In this section, we examine quality metrics described in Section~\ref{sxn:methods} for several NLP model architectures.
%
%In particular, 
Within the past two years, nearly 100 open source, pre-trained NLP DNNs based on the revolutionary Transformer architecture have emerged.
These include variants of BERT, Transformer-XML, GPT, etc.
%
The Transformer architectures consist of blocks of so-called Attention layers, containing two large, Feed Forward (Linear) weight matrics~\cite{Attn2017}. 
In contrast to smaller pre-Activation maps arising in Cond2D layers, Attention matrices are significantly larger.
In general, we have found that they have larger $\alpha$ PL exponents.
Based on HT-SR Theory (in particular, the interpretation of values of $\alpha \sim 2$ as modeling systems with good correlations over many size scales~\cite{BouchaudPotters03, SornetteBook}), this suggests that these models fail to capture successfully many of the correlations in the data (relative to their size) and thus that they are substantially \emph{under}-trained.
%
More generally, these NLP models are quite different than the CV models we considered in Section~\ref{sxn:cv}, both in terms of their architecture (e.g., wheter there are convolutions) as well as how they are used (e.g., whether one uses prediction error or perplexity).
Thus, they provide a very different test for our metrics.

\michael{WHERE: Norm-base metrics perform reasonably well on good NLP models, but they have problems with less well-trained models.
PL-based metrics (the Log $\alpha$-Norm metric, $\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$, and the Weighted Alpha metric, $\hat\alpha =\alpha\log\lambda_{max} $) do even better.
We highlight how to use the theory to identify poorly trained models where the empirical norm metrics will perform poorly.}

\michael{WHERE: We show that this not only works for CV, so we look at NLP.
We now look in detail at the alpha metric, the spectral norm, and the new norm.
We compare good and bad models, and a series of increasingly big models
trained on big data sets.  We show why alpha works but spectral norm is not enough.
The layer Spectral Norm behaves unexpectedly, while $\alpha$-based metrics can give insight into how well trained a model is.}

\michael{WHERE: Say large $\alpha$ values don't mean much.}

\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}

@@

\charles{More details here.}
The GPT models we analyze are deployed with the popular HuggingFace PyTorch library~\cite{XXX-XXX}.
GPT has 12 layers, with 4 Multi-head Attention Blocks, giving $48$ Layer Weight Matrices $\mathbf{W}$.
Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
The projection layer concatenates the self-attention results into a vector (of dimension $768$).
This gives $50$ large matrices.
%
Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
GPT  has an initial Token and Positional Embedding layers, of dimension $(40478\times 768)$ and $(512\times 768)$, respectively, whereas GPT2 has input Embeddings of shape $(50257\times 768)$ and $(1024\times 768)$, respectively. 
%
The OpenAI GPT2 (English) models are: \nred{gpt-small, gpt-medium, gpt-large, and gpt-xl}, \
having include $12, 24, 36, \text{and }48$ layers, respectively, with increasingly larger weight matrices.
The model card for GPT2 is published on github.\footnote{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}.
Table \ref{table:nlp} reports results for the average log norm metrics and the Weighted Alpha metric, 
\michael{Fix that.}

@@

\begin{table}[t]
\small
\begin{center}
%\begin{tabular}{|p{1in}|c|c|c|c|c|}
\begin{tabular}{|p{0.75in}|c|c|c|c|c|}
\hline
%   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
% Series & \#Layers   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
        &        & Log                   & Log                        & Weighted       & Log                                 \\
        & Num.   & $\Vert\cdot\Vert_{F}$ & $\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
 Series & Layers & Metric                & Metric                     & Metric         & Metric                              \\
\hline
GPT & 49 & 1.64  & 1.72 & 7.01 & 7.28 \\
GPT (small) & 49 & 2.04  & 2.54& 9.62 & 9.87 \\
GPT2 medium & 98 & 2.08 & 2.58& 9.74 & 10.01 \\
GPT2 large & 146 & 1.85 & 1.99& 7.67 & 7.94 \\
GPT2 xl & 194 & 1.86 & 1.92 & 7.17 & 7.51 \\
\hline
\end{tabular}
\end{center}
\caption{Average value for each of the four metrics for pretrainnd OpenAI GPT and GPT2 models.}
\label{table:nlp}
\end{table}
